---
title: "CIS_Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CIS_Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
Load Package: 
```{r}
library(dplyr)
source("R/filter_snps.R")
source("R/overlap_snps.R")
source("R/coloc_pipeline.R")
```

#Run Pipeline for a Single Cell Type
Load Gwas file:
```{r}
gwas <- read.table("/Users/tarrindewberry/Downloads/Gwas_raw_sumstats.tsv.gz", header = TRUE)
head(gwas)
```
Load eqtl file for a specific immune cell type:
In this example we will use a bcell_eqtl
```{r}
bcell_eqtl <- read.table("/Users/tarrindewberry/Downloads/bcell.cc.tsv.gz", header = TRUE)
head(bcell_eqtl)
```
#Filter the full genome raw statistics files searching for the given SNP that you are looking for. 
```{r}
gwas_filter_snps <- filter_nearby_snps_gwas(df_gwas =gwas,snp_id = "rs2453555", region_window = 50000)
bcell_filter_snps <- filter_nearby_snps_eqtl(df_eqtl = bcell_eqtl,snp_id = "rs2453555",region_window = 50000)
```
#Overlap the dataframes to ensure that the same snps are in both the eQTL and Gwas files.
```{r}
overlap_gwas_bcell <- overlap_snps(gwas_filter_snps,bcell_filter_snps)
```
#This what the overlap dataframe looks like
```{r}
head(overlap_gwas_bcell)
```
#We have to use this overlap dataframe to run the colocalization pipeline by creating, create two input list for the coloc functions
#Now we can run the coloc pipeline:
```{r}
run_coloc_pipeline(overlap_gwas_bcell, "bcell")
```
#The coloc function produces a colocalization percentage of 35%, showing that 

#Run Pipeline for Multiple Immune Cell Type: 
#For Bcell, Cd4 naive, and Cd8 naive 
```{r}
gwas <- read.table("/Users/tarrindewberry/Downloads/Gwas_raw_sumstats.tsv.gz", header = TRUE)
head(gwas)
```
#Create Load all Cell types
```{r}
cell_type_list <- list(
  "bcell" = "/Users/tarrindewberry/Downloads/bcell.cc.tsv.gz",
  "cd4_naive" = "/Users/tarrindewberry/Downloads/cd4_naiv.cc.tsv.gz",
  "cd8_naive" = "/Users/tarrindewberry/Downloads/cd8_naiv.cc.tsv.gz"
)

for(cell in names(cell_type_list)){
  cell_eqtl <- read.table(cell_type_list[["cell"]], header = TRUE)
  filter_nearby_snps_eqtl(cell_eqtl)
  #save as an object (load it)
}

```

```{r setup}
library(CISpackage)
```
